\section{Nodes}
Each node uses the information coming from a topic or from a device connected to the computer and publishes its results in another topic, which is then used by one or more nodes. The functionality of each node is described below.

\subsection*{\lstinline[style=cinline]{/lli_node}}
This node is in charge of reading the messages that come from the LLI and publishing them in the \lstinline[style=cinline]{/samples} topic. It is also responsible of sending the commands, which are published in the \lstinline[style=cinline]{/lli_input} topic, to the LLI. These commands need to be coded in a message, such that the LLI can read them.

\subsection*{\lstinline[style=cinline]{/sensor_node}}
The purpose of this node is to decode the information of the IMU that is packed in the messages of the \lstinline[style=cinline]{/samples} topic. This is done extracting the data from a string and converting it in the measurements of the accelerometer, the magnetometer and the gyroscope by transforming them into the correct units. This information is then published in the \lstinline[style=cinline]{/imu} topic.

\subsection*{\lstinline[style=cinline]{/gps_node}}
This node has two main functionalities. 

It reads the position information that comes from the GPS from the same serial port and decodes it to know the latitude and longitude of the vessel. With this information it is able to compute the relative distance of the vessel with the chosen origin of the NED frame, given by its latitude and longitude. The latitude and longitude of the vessel, together with its relative position to the origin is published in the \lstinline[style=cinline]{/gps_pos} topic.

Additionally, it parses the correction data from the RTK base to the GPS in the vessel using the serial port. A more detailed description of the RTK base can be found in \autoref{app:rtk_gps}. 

\subsection*{\lstinline[style=cinline]{/KF_attitude_node}}
The attitude Kalman filter is implemented in this node using the information that comes from the \lstinline[style=cinline]{/imu} topic. This node uses that data to estimate the angular position, velocity and acceleration of the vessel as described in \autoref{sec:attFusion}. Finally, the estimation is published in the \lstinline[style=cinline]{/kf_attitude} topic.

\subsection*{\lstinline[style=cinline]{/KF_position_node}}
The estimation of the position of the vessel is done in this node, as described in \autoref{sec:posFusion}. The information of the GPS is fused with the measurements of the accelerometer and the estimated attitude to give a better estimate of the translational position, velocity and acceleration of the vessel. This estimation is then published in the \lstinline[style=cinline]{/kf_position} topic.

\subsection*{\lstinline[style=cinline]{/path_follower_node}}
This node implements the path follower algorithm described in \autoref{sec:pathfollower}. It reads the waypoints generated as described in \autoref{sec:pathgeneration} from a .txt file, and the estimated position and attitude from the Kalman filters. With this information it is able to compute the required heading reference for the vessel to reach the desired path, which is published in the \lstinline[style=cinline]{/control_reference} topic.

\subsection*{\lstinline[style=cinline]{/controller_node}}
The inner controller node uses the information from both filters as well as the reference published in the \lstinline[style=cinline]{/control_reference} topic to apply the gains, both state feedback and integral control, and computes the required force in each motor. These forces are finally translated to PWM values to be published in the \lstinline[style=cinline]{/lli_input} topic.

\subsection*{\lstinline[style=cinline]{/keyboard_teleop_node}}
This node send commands to the motors using the \lstinline[style=cinline]{/lli_input} topic. The commands depend on the keys that are pressed:
\begin{itemize}
    \item '1': Enable the commands.
    \item '0': Disable the commands and stop the motors.
    \item 'W'/'A': Increase/decrease the speed of the motors.
    \item 'S'/'D': Turn left/right. One of the motors rotates with half of its speed.
    \item 'H': Gives the motors a step in speed in the forward direction.
    \item 'T': Gives the motors a step in speed to turn counterclockwise.
\end{itemize}


