\chapter*{Preface}
\vspace{-12 pt}
The focus of this project is to design a control system for an autonomous surface vessel, such that it can navigate autonomously over a predetermined area in the water.

This report has been written by a group of students on the second semester of the Master in Control and Automation at Aalborg University in the spring semester of 2017. It has been supervised by Jesper Abildgaard Larsen, associate professor at the Institute of Electronic Systems at Aalborg University. 

The group would like to thank Jens Frederik Dalsgaard Nielsen for his guidance during the setup of the GPS, and Palle Andersen for his help in the robust control design, both associate professors at the Institute of Electronic Systems Aalborg University.

The reader is expected to have a basic knowledge within physics and mathematics, as well as in modeling and linear control theory.

\textbf{Reading Instructions}
\vspace{-10 pt}
\begin{itemize}
    \item[-] The report is divided in three parts. Part I deals with the analysis of the system, which includes a description of the setup and the derivation of the dynamic model. Part II includes the control design, which contains the two approaches for an inner controller, the outer controller, the sensor fusion and the implementation. Part III includes the results of the project, the discussion and the conclusion.
    \item[-] The report also includes appendixes that contain the journals for the different tests and other relevant information.
    \item[-] The bibliography is written using ISO 690, noted as [x], and it is included at the end of the report, after the appendixes.
    \item[-] An attachment is included as part of the report, and contains MATLAB scripts and simulations files, test data files and the ROS workspace files used in the project.
\end{itemize}

%
\textbf{Text by:}\\
\vspace{-12 pt}
\begin{table}[H]
	\centering
		\begin{tabular}{c c c}
			\underline{\phantom{JAERJAERJAERJAERGO}} & \phantom{cookies} & \underline{\phantom{JAERJAERJAERJAERGO}} \\
			Alejandro Alonso García & \phantom{cookies} & Anders Egelund Kjeldal \\
			&&\\
			\underline{\phantom{JAERJAERJAERJAERGO}} & \phantom{cookies} & \underline{\phantom{JAERJAERJAERJAERGO}} \\
			Himal Kooverjee & \phantom{cookies} & Niels Skov Vestergaard		\\
			&&\\
	    \multicolumn{3}{c}{\underline{\phantom{JAERJAERJAERJAERGO}}}\\
	    \multicolumn{3}{c}{Noelia Villarmarzo Arruñada}\\				
		\end{tabular}
\end{table}
\pagebreak